ABCD_Prelim_Analyses

### 2021-11-15 ABCD_Prelim_Analyses.Rmd
Updated files

# calculate a total FPN score
combined.data.noOutliers$totalFPN <- rowSums(data.frame(as.numeric(combined.data.noOutliers$df4),as.numeric(combined.data.noOutliers$df16),as.numeric(combined.data.noOutliers$df18)))
```

@@

# Replicate findings from Cui et al., 2020
summary(lm(thompson_PC2~interview_age+sex+df18,data=combined.data.noOutliers)) #0.06
summary(lm(df18~interview_age,data=combined.data.noOutliers)) # **
summary(lm(thompson_PC2~interview_age+sex+df16,data=combined.data.noOutliers))
summary(lm(df16~interview_age,data=combined.data.noOutliers)) # **
summary(lm(thompson_PC2~interview_age+sex+df4,data=combined.data.noOutliers))
summary(lm(df4~interview_age,data=combined.data.noOutliers)) # **

# See which networks are associated with which cognitive factor scores
summary(lm(thompson_PC1~interview_age+sex+df2+df3+df4+df5+df6+df7+df8+df9+df10+df11+df12+df13+df14+df15+df16+df17+df18,data=combined.data.noOutliers))
summary(lm(thompson_PC2~interview_age+sex+df2+df3+df4+df5+df6+df7+df8+df9+df10+df11+df12+df13+df14+df15+df16+df17+df18,data=combined.data.noOutliers))
summary(lm(thompson_PC3~interview_age+sex+df2+df3+df4+df5+df6+df7+df8+df9+df10+df11+df12+df13+df14+df15+df16+df17+df18,data=combined.data.noOutliers))

summary(lm(df4~interview_age+sex+abuse,data=combined.data.noOutliers))
summary(lm(df4~interview_age+sex+totalACEs,data=combined.data.noOutliers))

summary(lm(df16~interview_age+sex+abuse,data=combined.data.noOutliers))
summary(lm(df16~interview_age+sex+totalACEs,data=combined.data.noOutliers))

summary(lm(df18~interview_age+sex+abuse,data=combined.data.noOutliers))
summary(lm(df18~interview_age+sex+totalACEs,data=combined.data.noOutliers))

summary(lm(thompson_PC1~interview_age+sex+totalACEs,data=combined.data.noOutliers)) # ***
summary(lm(thompson_PC2~interview_age+sex+totalACEs,data=combined.data.noOutliers)) # *
summary(lm(thompson_PC3~interview_age+sex+totalACEs,data=combined.data.noOutliers)) # ***
summary(lm(thompson_PC1~interview_age+sex+abuse,data=combined.data.noOutliers)) # *
summary(lm(thompson_PC2~interview_age+sex+abuse,data=combined.data.noOutliers)) #ns
summary(lm(thompson_PC3~interview_age+sex+abuse,data=combined.data.noOutliers)) # *

summary(lm(thompson_PC1~interview_age+sex+df18,data=combined.data.noOutliers)) # ***
summary(lm(thompson_PC1~interview_age+sex+df16,data=combined.data.noOutliers)) # ***
summary(lm(thompson_PC1~interview_age+sex+df4,data=combined.data.noOutliers)) # ns

# total FPN:
summary(lm(thompson_PC1~interview_age+sex+totalFPN,data=combined.data.noOutliers)) # ***
summary(lm(thompson_PC2~interview_age+sex+totalFPN,data=combined.data.noOutliers)) # .
summary(lm(thompson_PC3~interview_age+sex+totalFPN,data=combined.data.noOutliers)) # .

summary(lm(thompson_PC1~interview_age+sex+totalFPN,data=combined.data.noOutliers[combined.data.noOutliers$totalACEs==0,]))
summary(lm(thompson_PC1~interview_age+sex+totalFPN,data=combined.data.noOutliers[combined.data.noOutliers$totalACEs>0,]))

ggplot(data=combined.data.noOutliers,aes(totalFPN,thompson_PC1))+geom_point()+geom_smooth(method="lm")
ggplot(data=combined.data.noOutliers,aes(totalFPN,thompson_PC2))+geom_point()+geom_smooth(method="lm")
ggplot(data=combined.data.noOutliers,aes(totalFPN,thompson_PC3))+geom_point()+geom_smooth(method="lm")

combined.data.noOutliers$ACEsFactor <- factor(combined.data.noOutliers$totalACEs==0,labels = c("ACEs","No ACEs"))
ggplot(data=combined.data.noOutliers,aes(totalFPN,thompson_PC1,color=ACEsFactor))+geom_point(size=1)+geom_smooth(method="lm")

@@

ggplot(data=combined.data.noOutliers,aes(interview_age,df18,color=ACEsFactor))+geom_point(size=1)+geom_smooth(method="lm")

ggplot(data=combined.data.noOutliers,aes(interview_age,totalFPN,color=ACEsFactor))+geom_point(size=1)+geom_smooth(method="lm")
```

>>>>

# calculate a total FPN score
combined.data.noOutliers$totalFPN <- rowSums(data.frame(as.numeric(combined.data.noOutliers$df4),as.numeric(combined.data.noOutliers$df16),as.numeric(combined.data.noOutliers$df18)))

# separate train and test sets
trainIDs <- read.csv("/Users/ariellekeller/Documents/Kellernet_PrelimAnalysis/train.csv")
train.data <- merge(combined.data.noOutliers,trainIDs,by="subjectkey") # n=2,143
test.data <- combined.data.noOutliers[!is.element(combined.data.noOutliers$subjectkey,trainIDs$subjectkey),] #n=2,424
```

@@

# Try to replicate findings from Cui et al., 2020
summary(lm(thompson_PC2~interview_age+sex+df18,data=train.data))
summary(lm(df18~interview_age,data=train.data)) # **
summary(lm(thompson_PC2~interview_age+sex+df16,data=train.data))
summary(lm(df16~interview_age,data=train.data)) # **
summary(lm(thompson_PC2~interview_age+sex+df4,data=train.data))
summary(lm(df4~interview_age,data=train.data)) # **

# See which networks are associated with which cognitive factor scores
summary(lm(thompson_PC1~interview_age+sex+df2+df3+df4+df5+df6+df7+df8+df9+df10+df11+df12+df13+df14+df15+df16+df17+df18,data=train.data))
summary(lm(thompson_PC2~interview_age+sex+df2+df3+df4+df5+df6+df7+df8+df9+df10+df11+df12+df13+df14+df15+df16+df17+df18,data=train.data))
summary(lm(thompson_PC3~interview_age+sex+df2+df3+df4+df5+df6+df7+df8+df9+df10+df11+df12+df13+df14+df15+df16+df17+df18,data=train.data))

summary(lm(df4~interview_age+sex+abuse,data=train.data))
summary(lm(df4~interview_age+sex+totalACEs,data=train.data))

summary(lm(df16~interview_age+sex+abuse,data=train.data))
summary(lm(df16~interview_age+sex+totalACEs,data=train.data))

summary(lm(df18~interview_age+sex+abuse,data=train.data))
summary(lm(df18~interview_age+sex+totalACEs,data=train.data))

summary(lm(thompson_PC1~interview_age+sex+totalACEs,data=train.data)) # ***
summary(lm(thompson_PC2~interview_age+sex+totalACEs,data=train.data)) # .
summary(lm(thompson_PC3~interview_age+sex+totalACEs,data=train.data)) # ***
summary(lm(thompson_PC1~interview_age+sex+abuse,data=train.data)) #ns
summary(lm(thompson_PC2~interview_age+sex+abuse,data=train.data)) #ns
summary(lm(thompson_PC3~interview_age+sex+abuse,data=train.data)) #ns

summary(lm(thompson_PC1~interview_age+sex+df18,data=train.data)) # ***
summary(lm(thompson_PC1~interview_age+sex+df16,data=train.data)) # ***
summary(lm(thompson_PC1~interview_age+sex+df4,data=train.data)) # ns

# total FPN:
summary(lm(thompson_PC1~interview_age+sex+totalFPN,data=train.data)) # **
summary(lm(thompson_PC2~interview_age+sex+totalFPN,data=train.data)) # 
summary(lm(thompson_PC3~interview_age+sex+totalFPN,data=train.data)) #

summary(lm(thompson_PC1~interview_age+sex+totalFPN,data=train.data[train.data$totalACEs==0,]))
summary(lm(thompson_PC1~interview_age+sex+totalFPN,data=train.data[train.data$totalACEs>0,]))

ggplot(data=train.data,aes(totalFPN,thompson_PC1))+geom_point()+geom_smooth(method="lm")
ggplot(data=train.data,aes(totalFPN,thompson_PC2))+geom_point()+geom_smooth(method="lm")
ggplot(data=train.data,aes(totalFPN,thompson_PC3))+geom_point()+geom_smooth(method="lm")

train.data$ACEsFactor <- factor(train.data$totalACEs==0,labels = c("ACEs","No ACEs"))
test.data$ACEsFactor <- factor(test.data$totalACEs==0,labels = c("ACEs","No ACEs"))

ggplot(data=train.data,aes(totalFPN,thompson_PC1,color=ACEsFactor))+geom_point(size=1)+geom_smooth(method="lm")

@@

ggplot(data=train.data,aes(interview_age,df18,color=ACEsFactor))+geom_point(size=1)+geom_smooth(method="lm")

ggplot(data=train.data,aes(interview_age,totalFPN,color=ACEsFactor))+geom_point(size=1)+geom_smooth(method="lm")

summary(lm(thompson_PC1~interview_age+sex+totalFPN,data=combined.data.noOutliers)) # path B ***
summary(lm(thompson_PC1~interview_age+sex+ACEsFactor,data=combined.data.noOutliers)) # path C ***
summary(lm(totalFPN~interview_age+sex+ACEsFactor,data=combined.data.noOutliers)) # path A ns
summary(lm(thompson_PC1~interview_age+sex+totalFPN+ACEsFactor,data=combined.data.noOutliers)) # path C' *** ***

summary(lm(thompson_PC1~interview_age+sex+df18,data=combined.data.noOutliers)) # path B ***
summary(lm(thompson_PC1~interview_age+sex+ACEsFactor,data=combined.data.noOutliers)) # path C ***
summary(lm(df18~interview_age+sex+ACEsFactor,data=combined.data.noOutliers)) # path A ns
summary(lm(thompson_PC1~interview_age+sex+df18+ACEsFactor,data=combined.data.noOutliers)) # path C' *** ***

summary(lm(thompson_PC1~interview_age+sex+df16,data=combined.data.noOutliers)) # path B ***
summary(lm(thompson_PC1~interview_age+sex+ACEsFactor,data=combined.data.noOutliers)) # path C ***
summary(lm(df16~interview_age+sex+ACEsFactor,data=combined.data.noOutliers)) # path A ns
summary(lm(thompson_PC1~interview_age+sex+df16+ACEsFactor,data=combined.data.noOutliers)) # path C' *** ***

summary(lm(thompson_PC1~interview_age+sex+totalFPN*ACEsFactor*sex,data=combined.data.noOutliers))

ggplot(data=train.data[train.data$sex=="M",],aes(ACEsFactor,totalFPN))+geom_col()
ggplot(data=train.data[train.data$sex=="F",],aes(ACEsFactor,totalFPN))+geom_bar()
```

#Plotting
```{r}

df.summary <- train.data %>%
     group_by(ACEsFactor,sex) %>%
     summarise(
         se = sd(totalFPN, na.rm = TRUE)/sqrt(length(totalFPN)),
         totalFPN = mean(totalFPN))
ggplot(data=df.summary, aes(x=sex,y=totalFPN,fill=as.factor(interaction(ACEsFactor,sex)))) +
  geom_bar(stat="identity",position="dodge") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +xlab("Biological Sex") + ylab("Total FPN (3, 15, 17) Topography") + geom_errorbar(aes(ymin=totalFPN-se, ymax=totalFPN+se), width=.2,position=position_dodge(.9))+coord_cartesian(ylim=c(11500,12000))

df.summary <- train.data %>%
     group_by(ACEsFactor,sex) %>%
     summarise(
         se = sd(thompson_PC1, na.rm = TRUE)/sqrt(length(thompson_PC1)),
         thompson_PC1 = mean(thompson_PC1))
ggplot(data=df.summary, aes(x=sex,y=thompson_PC1,fill=as.factor(interaction(ACEsFactor,sex)))) +
  geom_bar(stat="identity",position="dodge") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +xlab("Biological Sex") + ylab("Thompson PC1") + geom_errorbar(aes(ymin=thompson_PC1-se, ymax=thompson_PC1+se), width=.2,position=position_dodge(.9))+coord_cartesian(ylim=c(220,250))

df.summary <- train.data %>%
     group_by(ACEsFactor,sex) %>%
     summarise(
         se = sd(df16, na.rm = TRUE)/sqrt(length(df16)),
         df16 = mean(df16))
ggplot(data=df.summary, aes(x=sex,y=df16,fill=as.factor(interaction(ACEsFactor,sex)))) +
  geom_bar(stat="identity",position="dodge") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +xlab("Biological Sex") + ylab("Net 15 Topography") + geom_errorbar(aes(ymin=df16-se, ymax=df16+se), width=.2,position=position_dodge(.9))+coord_cartesian(ylim=c(1000,1500))

df.summary <- train.data %>%
     group_by(ACEsFactor,sex) %>%
     summarise(
         se = sd(df4, na.rm = TRUE)/sqrt(length(df4)),
         df4 = mean(df4))
ggplot(data=df.summary, aes(x=sex,y=df4,fill=as.factor(interaction(ACEsFactor,sex)))) +
  geom_bar(stat="identity",position="dodge") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +xlab("Biological Sex") + ylab("Net 3 Topography") + geom_errorbar(aes(ymin=df4-se, ymax=df4+se), width=.2,position=position_dodge(.9))+coord_cartesian(ylim=c(6000,6500))

df.summary <- train.data %>%
     group_by(ACEsFactor,sex) %>%
     summarise(
         se = sd(thompson_PC3, na.rm = TRUE)/sqrt(length(thompson_PC3)),
         thompson_PC3 = mean(thompson_PC3))
ggplot(data=df.summary, aes(x=sex,y=thompson_PC3,fill=as.factor(interaction(ACEsFactor,sex)))) +
  geom_bar(stat="identity",position="dodge") + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(), axis.line = element_line(colour = "black")) +xlab("Biological Sex") + ylab("Thompson PC3") + geom_errorbar(aes(ymin=thompson_PC3-se, ymax=thompson_PC3+se), width=.2,position=position_dodge(.9))+coord_cartesian(ylim=c(220,250))
